---
title: "Meta-learning how to forecast time series"
author: |
        | Thiyanga S. Talagala
        | Rob J Hyndman
        | George Athanasopoulos
date: "Monash University, Australia"
output: 
  beamer_presentation:
     theme: "Warsaw"
     includes:
        in_header: header_pagenrs.tex
---

```{r initial, echo = FALSE, cache = TRUE, include = FALSE}
library(knitr)
library(pander)
library(seer)
library(Mcomp)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
opts_chunk$set(fig.lp = '')
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.path = "figure/",
  cache.path = "cache/",
  cache = TRUE,
  fig.show = "hold",
  external = TRUE,
  dev = "png"
)

```

```{r set-options, echo=FALSE, cache=TRUE}
options(width = 50)
```


# Large collections of time series 

  \centerline{\includegraphics[width=\textwidth,height=7.5cm,keepaspectratio]{figures/example.jpg}}

- Forecasting demand for thousands of products across multiple
warehouses.


# Time series features

\begin{block}{Objective}
Develop a framework that automates the selection of the most appropriate forecasting model for a given time series by using an array of \textcolor{red}{features} computed from the time series.
\end{block}

\begin{itemize}[<+->]

\item \textbf{Basic idea:}
  
  Transform a given time series $y=\{y_1, y_2, \cdots, y_n\}$ to a feature vector $F = (f_1(y), f_2(y), \cdots, f_p(y))'$. 

\item Examples for time series features

\begin{itemize}[<+->]
\vfill\item strength of trend
\vfill\item strength of seasonality
\vfill\item lag-1 autocorrelation
\vfill\item spectral entropy

\end{itemize}

\end{itemize}


# Feature-space of time series

\begin{block}{STL-decomposition}
$Y_t=T_t+S_t+R_t$
\end{block}

\begin{itemize}
\vfill\item strength of trend: $1-\frac{Var(R_t)}{Var(Y_t-S_t)}$
\vfill\item strength of seasonality: $1-\frac{Var(R_t)}{Va(Y_t-T_t)}$
\end{itemize}

\begincols
\begincol{.48\textwidth}


```{r, warning=FALSE, message=F, echo=FALSE, fig.height=10}
library(Mcomp)
library(ggplot2)
library(grid)
library(gridExtra)
library(ggrepel)
library(png)
library(tsfeatures)
library(tidyverse)
library(ggpubr)

# Extract required series
series_id <- c("N0001", "N0633", "N0625", "N0645","N1912", "N2012")
six_series <- lapply(M3[series_id], function(u){u$x})
p <- lapply(six_series,
function(u) {autoplot(u) + xlab("") + ylab("")}
)
for (i in seq_along(six_series))
p[[i]] <- p[[i]] + ggtitle(series_id[i])+theme(title =element_text(size=30, face='bold'))
grid.arrange(grobs = p, ncol = 2)

```

\endcol \begincol{.48\textwidth}

\pause

```{r fig2, fig.height=10, message=F, echo=F, warning=F}
df <- tsfeatures(six_series, c("stl_features")) %>%
  select(trend, seasonal_strength) %>%
  rename(seasonality = seasonal_strength) %>%
  replace_na(list(seasonality = 0))
df$id <- names(six_series)
ggplot(df, aes(x = trend, y = seasonality)) +
  geom_point(size = 5, color = 'red') +
  xlim(0, 1) + ylim(0, 1) +
  coord_fixed() +
  geom_text_repel(
    aes(label = id),
    colour = "black",
    size = 10,
    box.padding = unit(0.5, "lines")
  ) +
  theme(legend.position = "none")
```

  \endcol
\endcols

# Time series features



\begin{columns}
\begin{column}{0.48\textwidth}
   \begin{itemize}
   \item length
    \item strength of seasonality
    \item strength of trend
    \item linearity
    \item curvature
    \item spikiness
    \item stability
    \item lumpiness
    \item first ACF value of remainder series
    \item parameter estimates of Holt's linear trend method
    \end{itemize} 
\end{column}
\begin{column}{0.48\textwidth}
   \begin{itemize}
    \item spectral entropy
    \item Hurst exponent
    \item nonlinearity
    \item parameter estimates of Holt-Winters' additive method
    \item unit root test statistics
    \item first ACF value of residual series of linear trend model 
    \item ACF and PACF based features - calculated on both the raw and differenced series
    \end{itemize} 
\end{column}
\end{columns}

# Methodology: FFORMS

**FFORMS:** **F**eature-based **FOR**ecast **M**odel **S**election

\textcolor{red}{Offline}: Classification algorithm is trained

\textcolor{red}{Online}: Use the classification algorithm to select appropriate forecast-models for new time series



# FFORMS: population

 \centerline{\includegraphics[width=\paperwidth]{figures/fw1.png}}

# FFORMS: observed sample

 \centerline{\includegraphics[width=\paperwidth]{figures/fw2.png}}
 
# FFORMS: simulated time series

 \centerline{\includegraphics[width=\paperwidth]{figures/fw3.png}}
 
# FFORMS: reference set
 
 \centerline{\includegraphics[width=\paperwidth]{figures/fw4.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw5.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw6.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw7.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw8.png}}


# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw9.png}}

# FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{figures/fw10.png}}

# FFOMS: Random-forest classifier

\centerline{\includegraphics[width=\paperwidth]{figures/fw11.png}}


# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw12.png}}


# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw13.png}}

# FFORMS: "online" part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{figures/fw14.png}}

# Application to M competition data

- Proposed algorithm is applied to yearly, quarterly and monthly series separately

- We run two experiments for each case.

\begin{table}[!htp]
\centering\footnotesize\tabcolsep=0.2cm
\def\yes{$\checkmark$}
\captionsetup{labelformat=empty}
\label{tbl:Mcomps}
\begin{tabular}{lccccr@{\hspace*{0.3cm}}cccc}
\toprule
                &  \multicolumn{ 4}{c}{Experiment 1} & & \multicolumn{ 4}{c}{Experiment 2} \\
                &    Source  &    Y  & Q  &    M &            &    Source  &    Y  & Q  &    M \\\cline{2-5}\cline{7-10}
& & & & & & & & & \\[-0.25cm]
Observed series &         M1 &        181 &        203 &        617 &            &         M3 &        645 &        756 &       1428 \\
     New series &         M3 &        645 &        756 &       1428 &            &         M1 &        181 &        203 &        617 \\
\bottomrule
\end{tabular}
\end{table}



# Discussion and Conclusions

\begin{itemize}[<+->]

\vfill\item Proposed a novel framework for forecast-model selection using meta-learning based on time series features.

\vfill\item Our method almost always performs better than common benchmark methods, and better than the best-performing methods from the M3 competition.

\vfill\item The framework is general and can be applied to any large collection of time series.

\vfill\item Advantage: Not necessary to estimate several different models for the data and undertake an empirical evaluation of their forecast accuracy on a given time series.

\end{itemize}


# R package: seer

 \vspace{0.15cm}
\centerline{\includegraphics[width=\textwidth,height=3.5cm,keepaspectratio]{figures/seer.png}}
\vspace{0.3cm}
 \centerline{available at:}
 \vspace{0.15cm}
 \centerline{https://github.com/thiyangt/seer}